Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-10-20 astronum_2017/astronum-amrex.tex
reset the figures to the top
add plot

Our suite of application codes is built on the
\amrex\ block-structured adaptive mesh refinement (AMR) library.  The
basic programming model has \amrex\ managing the grid data-structures
and parallel communications, and it calls the computational kernels on
a patch-by-patch basis.  The core library is written in \cpp\ with 
computational kernels written in Fortran\footnote{Currently
\maestro\ is written in pure-Fortran, but will be
 ported to the updated \cpp\ \amrex\ framework this coming year.}
---this allows us to take
advantage of the strengths of both languages. \amrex\ supports subcycling
in time, which we use in \castro.
\begin{figure}[htb]
\centering
\includegraphics[width=0.48\linewidth]{ppsn-amr.png}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:amr_grid} Colormap of density and the associated AMR 
grid structure in a pulsational pair-instability supernova simulation.  
Image courtesy of Ken Chen \cite{chen2014two}.}
\end{minipage}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:maestro_scaling} \maestro\ strong scaling on NERSC
  edison and OLCF titan for a 3-d XRB problem.  Two different problem
  sizes are shown.  We see excellent strong scaling to high core counts
  for this problem.\vspace{2em}}
\end{minipage}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\linewidth]{sedov_scaling}
\includegraphics[width=0.48\linewidth]{wdmerger_scaling}
\caption{\label{fig:castro-scaling} (left) \castro\ strong scaling on
  OLCF titan for a pure hydro (Sedov w/ real EOS) problem.  The
  different colors represent different base resolutions.  We see
  excellent strong and weak scaling for the single level runs.  For
  the $1024^3$ run, we also ran with one level of refinement by a
  factor of 2 (triangles) or a factor of 4 (squares), and see good
  strong scaling.  The variability (shown by the error bars) at high
  core counts shows we are becoming work-starved. (right)
  \castro\ strong scaling on OLCF titan for a hydro + Poisson gravity
  (wdmerger) problem.  This problem uses a multipole solver to
  determine Dirichlet boundary conditions representing an isolated
  mass distribution, and then geometric multigrid to solve for the
  potential in the interior.  Two coarse grid sizes are shown, and
  demonstrate great strong scaling.  We also look at a single level of
  refinement (by a factor of 4) on top of this coarse grid.}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\linewidth]{summitdev_scaling}
\includegraphics[width=0.48\linewidth]{cori_scaling}
\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new
  architectures.  The left figure shows the \castro\ GPU proxy app
  (\starlord) on OLCF summitdev.  We see good scaling behavior to 64
  GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
right, we explore performance on NERSC cori (and Intel KNL platform).}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:cudaode} GPU speed up over single CPU core for two
  reaction networks.  The 13-isotope alpha-chain is a
  potential network for the sub-Chandra simulations.
  The 9-isotope Urca network is used for our Urca simulations.  A
  variety of grid sizes were used, $16^3$, $32^3$, $64^3$, and
  $128^3$.  In all cases, we see significant speed up on the GPU
  vs.\ a single core.  These speed-up numbers are from our testbed
  workstation with a Titan X Pascal board and Intel Xeon X5650 processors.}
\end{minipage}
\end{figure}

@@

\begin{figure}[htb]
\centering \includegraphics[height=1.8in]{subch_h}
\includegraphics[height=1.8in]{xrb_compact}
\includegraphics[height=1.8in]{wdmerger_08030_new}
\caption{\label{fig:current-runs} (left) Convective plumes in a
  \maestro\ sub-Ch calculation. (center) Vertical velocity showing the
  convective structure in a \maestro\ XRB calculation. (right)
  Snapshot of a \castro\ simulation of the merger of two white dwarfs,
  with 0.90 and 0.81 solar masses. The contours represent density
  levels.}
\end{figure}

>>>>

Our suite of application codes is built on the
\amrex\ block-structured adaptive mesh refinement (AMR) library.  The
basic programming model has \amrex\ managing the grid data-structures
and parallel communications, and it calls the computational kernels on
a patch-by-patch basis.  The core library is written in \cpp\ with 
computational kernels written in Fortran\footnote{Currently
\maestro\ is written in pure-Fortran, but will be
 ported to the updated \cpp\ \amrex\ framework this coming year.}
---this allows us to take
advantage of the strengths of both languages. \amrex\ supports subcycling
in time, which we use in \castro.
\begin{figure}[t]
\centering
\includegraphics[width=0.58\linewidth]{slice_grid}
\begin{minipage}[b]{0.38\linewidth}
\caption{\label{fig:amr_grid} A slice through a
  \castro\ three-dimensional simulation of merging white dwarfs
  showing the 2-level grid structure.  The boxes making up the base
  grid are drawn in white and the boxes making up the finer grid are drawn in black.  Note
  that the finer boxes can overlap multiple coarse boxes.\vspace{2em}}
\end{minipage}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:maestro_scaling} \maestro\ strong scaling on NERSC
  edison and OLCF titan for a 3-d XRB problem.  Two different problem
  sizes are shown.  We see excellent strong scaling to high core counts
  for this problem.\vspace{2em}}
\end{minipage}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{sedov_scaling}
\includegraphics[width=0.48\linewidth]{wdmerger_scaling}
\caption{\label{fig:castro-scaling} (left) \castro\ strong scaling on
  OLCF titan for a pure hydro (Sedov w/ real EOS) problem.  The
  different colors represent different base resolutions.  We see
  excellent strong and weak scaling for the single level runs.  For
  the $1024^3$ run, we also ran with one level of refinement by a
  factor of 2 (triangles) or a factor of 4 (squares), and see good
  strong scaling.  The variability (shown by the error bars) at high
  core counts shows we are becoming work-starved. (right)
  \castro\ strong scaling on OLCF titan for a hydro + Poisson gravity
  (wdmerger) problem.  This problem uses a multipole solver to
  determine Dirichlet boundary conditions representing an isolated
  mass distribution, and then geometric multigrid to solve for the
  potential in the interior.  Two coarse grid sizes are shown, and
  demonstrate great strong scaling.  We also look at a single level of
  refinement (by a factor of 4) on top of this coarse grid.}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{summitdev_scaling}
\includegraphics[width=0.48\linewidth]{cori_scaling}
\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new
  architectures.  The left figure shows the \castro\ GPU proxy app
  (\starlord) on OLCF summitdev.  We see good scaling behavior to 64
  GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
right, we explore performance on NERSC cori (and Intel KNL platform).}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:cudaode} GPU speed up over single CPU core for two
  reaction networks.  The 13-isotope alpha-chain is a
  potential network for the sub-Chandra simulations.
  The 9-isotope Urca network is used for our Urca simulations.  A
  variety of grid sizes were used, $16^3$, $32^3$, $64^3$, and
  $128^3$.  In all cases, we see significant speed up on the GPU
  vs.\ a single core.  These speed-up numbers are from our testbed
  workstation with a Titan X Pascal board and Intel Xeon X5650 processors.}
\end{minipage}
\end{figure}

@@

\begin{figure}[t]
\centering \includegraphics[height=1.8in]{subch_h}
\includegraphics[height=1.8in]{xrb_compact}
\includegraphics[height=1.8in]{wdmerger_08030_new}
\caption{\label{fig:current-runs} (left) Convective plumes in a
  \maestro\ sub-Ch calculation. (center) Vertical velocity showing the
  convective structure in a \maestro\ XRB calculation. (right)
  Snapshot of a \castro\ simulation of the merger of two white dwarfs,
  with 0.90 and 0.81 solar masses. The contours represent density
  levels.}
\end{figure}
